Supplementary Materials for Imminent range shifts in transboundary fish stocks challenge their management under climate change

### 2021-05-11 Scripts/Supplements.Rmd
Removed TI direcction statement since it is obsolete

For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species. Moreover, it projects the future habitat suitability in each grid-cell (*i*) based on the following components;

@@

\begin{equation}
\begin{split}
\label{eq:one}
lat_{ts} = \mu (lat_{per}) \\
lon_{ts} = \mu (lon_{per})
\end{split}
\end{equation}

@@

\begin{equation}
\label{eq:two}
D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
\end{equation}

@@

\begin{equation}
\label{eq:three}
TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
\end{equation}

where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$. Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels. From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries. In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries. The index was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg].

@@

\begin{equation}
\label{eq:four}
ToE = TI_\mu[t']>2* TI_\sigma 
\end{equation}

@@

\begin{equation}
\label{eq:five}
\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
\end{equation}

@@

```{r toeFigureS6, eval = T,  echo = F, fig.width=14, fig.height=10, fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring Exclusive Economic Zones sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel)."}

img <- readPNG("../Figures/FigS6.png")
grid.raster(img)

>>>>

<!-- Included more details of the DBEM in response to reviewer 3 -->
For each species, the DBEM uses the Sea Around Us (www.searaundus.org) distributional data and environmental variables from 1970 to 2000 to estimate the current distribution and environmental profile of each species. Moreover, it projects the future habitat suitability in each grid-cell (*i*) based on the following components;

@@

\begin{equation}
\begin{split}
\label{eq:four}
lat_{ts} = \mu (lat_{per}) \\
lon_{ts} = \mu (lon_{per})
\end{split}
\end{equation}

@@

\begin{equation}
\label{eq:five}
D_{ens} = acos(sin(lat_{eez})*sin(lat_{ts}) + cos(lat_{eez})*cos(lat_{ts})*cos(lon_{eez}-lon_{ts}))
\end{equation}

@@

\begin{equation}
\label{eq:six}
TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
\end{equation}

where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$.

<!-- Removed in response to reviewer 3 -->
<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels. From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries. In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries. -->

The index was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg].

@@

\begin{equation}
\label{eq:seven}
ToE = TI_\mu[t']>2* TI_\sigma 
\end{equation}

@@

\begin{equation}
\label{eq:eight}
\Delta{SSR_{e,s}} = \frac{(SSR_{f}-SSR_{th})}{SSR_{th}}* 100
\end{equation}

@@
